Wavelet Multiresolution Spatio-Temporal Modelling Using the Integro-Difference Equation

### 2011-02-22 Multiresolution_framework/ltx/manuscript.tex
added state space representation of the model

@@

	T_s \varsigma\int_\Omega { 

@@

observation equation is given by 

@@

\parham{Mike you can add state-space derivation using Galerkin projection here.}
% however it can be tailored to capture underlying dynamics with desired accuracy for a sufficiently large value of $j$, a value between zero and three is usually sufficient in many applications \cite{Wei2006}. Note that number of translation parameters is always finite and depends on the range of measured data.

The state-space representation of the spatio-temporal IDE proposed in \cite{Dewar2009} is derived for an arbitrary set of basis functions and hence this derivation can be applied using scaling and wavelet basis functions. Let
\begin{equation}
 \mathbf{\Lambda}_{x}=\int_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\boldsymbol\mu\left(\mathbf{s}\right)^\top d\mathbf{s}
\label{eq:ConstructionIntegral1}
\end{equation}
\begin{equation}
 \mathbf{\Lambda}_{\theta}=\iint\limits_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\boldsymbol\theta^\top \boldsymbol\lambda\left(\mathbf{s-r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top d\mathbf{r}d\mathbf{s}
\label{eq:ConstructionIntegral2}
\end{equation}
where $\boldsymbol\Lambda_{x}, \boldsymbol\Lambda_{\theta} \in \mathbb{R}^{n_x \times n_x}$. The spatial IDE model in (\ref{eq:ConvolutionIntegral}) and (\ref{eq:observation equation}) can be represented as linear state-space model of the form  
\begin{equation}
 \mathbf{x}_{t+1}=\mathbf{A}\left(\boldsymbol\theta\right)\mathbf{x}_t+\mathbf{w}_t
\label{eq:StateEquation}
\end{equation}
\begin{equation}
\mathbf{y}_t=\mathbf{C}\mathbf{x}_t+\mathbf{v}_t
\label{eq:ObservationEquation}
\end{equation}
where 
\begin{equation}
 \mathbf{A}\left(\boldsymbol\theta\right)=\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda_{\theta}}
\label{eq:TransitionMatrix}
\end{equation}
and  \\ \\ 
$\mathbf{C}=$
$\begin{pmatrix}
\boldsymbol\phi_{j_0}\left(\mathbf{s}_1\right) & \boldsymbol\phi_{j_0}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\phi_{j_0}\left(\mathbf{s}_{n_y}\right) \\
\boldsymbol\psi_{j_0}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j_0}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j_0}\left(\mathbf{s}_{n_y}\right) \\
\boldsymbol\psi_{j_0+1}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j_0+1}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j_0+1}\left(\mathbf{s}_{n_y}\right) \\
\vdots & \vdots & \ddots& \vdots \\
\boldsymbol\psi_{j}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j}\left(\mathbf{s}_{n_y}\right) \\
\end{pmatrix}^{\top}$ \\ \\
\\ \\ and where 
\begin{equation}
 \mathbf{w}_t=\mathbf{\Lambda}_{x}^{-1}\int_{\mathcal{S}}\boldsymbol\mu \left(\mathbf{s}\right)e_t\left(\mathbf{s}\right)d\mathbf{s}
\label{CovofField}
\end{equation}
 is a vector valued, zero-mean normally distributed white noise process with covariance given by (Lemma \ref{lemma:FieldCovariance}, Appendix)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\boldsymbol\Pi\mathbf{\Lambda}_{x}^{-\top}
\end{equation}
and
\begin{equation}
 \boldsymbol\Pi=\iint\limits_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\eta\left(\mathbf{s}-\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^{\top}d\mathbf{r}d\mathbf{s}
\label{eq:Pi}
\end{equation}
By substituting the  parameter vector $\boldsymbol\theta$, the  state vector $\mathbf{x_t}$, the vector of kernel basis functions $\boldsymbol\lambda$ and the vector of field basis functions $\boldsymbol\mu$ in (\ref{eq:ConstructionIntegral1}-\ref{eq:Pi}), the state-space representation of the multiresolution approximation of the IDE can be achieved. It should be noted that the complexity of the model is independent of the number of observation locations, providing more flexibility in handling high resolution data. %The model is also capable of being applicable to scenarios with irregular observation locations.

>>>>

\begin{equation}\label{eq:PiecewiseSigmoid} 

@@

@@

Note that the parameter $\varsigma$ is absorbed into the kernel for simplicity. Oservation equation is given by 

@@

To derive the state-space representation of the model we substitute decompositions \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} into equation \eqref{DiscreteTimeModel} and then cross-multiply the decomposed model by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the spatial domain, $\Omega$, giving
\begin{align}\label{eq:DecomposedModel2} 
	&\int_{\Omega} \boldsymbol\mu  \left(\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top  d\mathbf{r}\mathbf{x}_{t+1}= 
	\xi\int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top d\mathbf{r}\mathbf{x}_t \nonumber \\
	&+T_s \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r'}\right)
	    \boldsymbol\mu^\top\left(\mathbf{r'}\right)\mathbf{x}_t 
	\, d\mathbf{r}'d\mathbf{r}}
	+ \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}, 
\end{align}
Now defning 
\begin{equation}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} \triangleq \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\mu^\top\left(\mathbf{r}\right) d\mathbf{r}
\end{equation}
\begin{equation}\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta}\triangleq\int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r'}\right)
	    \boldsymbol\mu^\top\left(\mathbf{r'}\right)\mathbf{x}_t 
	\, d\mathbf{r}'d\mathbf{r}}
\end{equation}
and substituting these into \eqref{eq:DecomposedModel2} and cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ yields
\begin{equation}\label{eq:StateEquation}
 \mathbf x_{t+1}=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t
\end{equation}
where
\begin{equation}\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta)=(T_s\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I)
\end{equation}
and $\mathbf I$ is the identity matrix, and 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance \cite{Scerri2009}
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\eta\left(\mathbf{r}-\mathbf{r'}\right)\boldsymbol\mu^{\top}\left(\mathbf{r'}\right)d\mathbf{r'}d\mathbf{r}\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
Similarly, the observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is 
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\mu}_j(\mathbf{r}') \, d\mathbf{r}'.
\end{equation}

\parham{Mike you can add state-space derivation using Galerkin projection here.}
By substituting the  parameter vector $\boldsymbol\theta$, the  state vector $\mathbf{x_t}$, the vector of kernel basis functions $\boldsymbol\lambda$ and the vector of field basis functions $\boldsymbol\mu$ in (\ref{eq:ConstructionIntegral1}-\ref{eq:Pi}), the state-space representation of the multiresolution approximation of the IDE can be achieved. It should be noted that the complexity of the model is independent of the number of observation locations, providing more flexibility in handling high resolution data. %The model is also capable of being applicable to scenarios with irregular observation locations.
